---
title: "Project 1 - Text Mining Chess Data"
author: "Luis Calleja"
date: "September 19, 2016"
output: html_document
---

```{r, echo=FALSE}
library(stringr);library(plyr);library(ggplot2)
```
In this project, the Output goal is a dataframe with the following informaiton (in order): player's name, player's state, total # of points, player's pre-rating, avg pre chess rating of opponents.
```{r, warning=FALSE}
g<-file('/home/lechuza/Documents/CUNY/data_class/project1/tournamentinfo.txt')
raw<-readLines(g)
```
ReadLines (above) splits on line breaks; however, each player spans two lines. In order to get the desired output of one line per player, we'll apply regex to concatenated elements and split on "-{3,}" (dashes).

```{r}
#raw_c<-str_c(raw[-(0:4)])
raw_c<-paste(raw[-(0:4)], collapse = '')
```


Investigate that the collapse did not result in unintended results
```{r, eval=FALSE}
#ouput supressed
nchar(raw_c)
g<-sapply(raw,nchar)
sum(g)
```
The lengths/number of characters are nearly the same, so we are good.

```{r}
raw_split<-unlist(str_split(raw_c,"-{3,}"))
```
Commence string extraction; the convenient feature is that 
we can do our computations at the local level, then apply 
it to the vector elements as a vector function

###Text extraction
We start with name.

We know that names appear in the second slot, so we could split
on the pipe "|" then run the regex.

```{r, eval=FALSE, echo=FALSE}
# a test
raw_split[3]
#let's first split on the pipe "|"... have to use escape syntax
check<-unlist(str_split(raw_split[3],"\\|"))
```
Let's run this string split on the entire vector to determine
how normalized is the data. We will run some diagnostics, ie ensure equal number of elements created from the split.

```{r}
spl.1<-function(x) {
  unlist(str_split(x,"\\|"))
}
te.1<-sapply(raw_split,spl.1,USE.NAMES = FALSE)
te.1<-te.1[-65]
```

```{r, eval=FALSE}
#output suppressed
#ensure equal number of elements for all vectors
gf<-sapply(te.1,function(x) length(unlist(x)))
any(gf!=21)
```
The above proves that each player field has 21 elements. Hence the data is normalized and can be placed in a structured object.

```{r}
h<-data.frame(sapply(te.1,function(x) rbind(unlist(x))),stringsAsFactors = FALSE)
```

Let's clean up each element in the dataframe and transpose it.
```{r, echo=FALSE, eval=FALSE}
#originally whigged because colwise() wasn't working properly, but then it started working...
#str_trim is a vectorized function

#doesn't work... returns levels
h.1<-data.frame(sapply(h,MARGIN=2,str_trim,side="both"))

rig<-function(x){
  as.character(str_trim(x))
}
#doesn't work because returns levels
h.2<-data.frame(apply(h,MARGIN=2,rig))
```

```{r}
#works as intended
#dr<-as.character(str_trim(h[,4]))
h.1<-colwise(str_trim)(h)
#transpose the dataframe
h.t<-as.data.frame(t(h.1),stringsAsFactors = FALSE)
#convert all these factors to characters

#put column names on the dataframe...
```

We can start to build the 'production' dataframe. First we create one with the player's name and state:
```{r}
#h.t[1:5,3]
final<-h.t[,c(2,11,3)]
names(final)<-c('name','state','total.pts')
head(final)
```

Extract pre-rating from 12th column.
```{r, echo=FALSE, eval=FALSE}
# 1436P23
h.t[61,12]
str_sub(b,0,-2)
str_trim(unlist(str_extract_all(h.t[39,12],"([[:digit:]]){2,}P.+?(?=-)")))
#http://stackoverflow.com/questions/7124778/how-to-match-anything-up-until-this-sequence-of-characters-in-a-regular-expres

#combine the above with 
str_trim(unlist(str_extract_all(x," ([[:digit:]]){2,} ")))
```

There are a couple of iterations on how the pre-rating appears in the data, we combine regex to mine for both with an 'OR' operation.
```{r}
#h.t[39,12]
#str_trim(unlist(str_extract_all(h.t[39,12]," ([[:digit:]]){2,} |([[:digit:]]){2,}P.+?(?=-)")))

#we'll run two operations... first the initial extract, then we'll remove all p
xtract.prerating<-function(x){
  str_trim(unlist(str_extract_all(x," ([[:digit:]]){2,} |([[:digit:]]){2,}P.+?(?=-)")))
}

j<-sapply(h.t[,12],xtract.prerating)

dig<-function(x) {unlist(str_extract_all(x,"[[:digit:]]{3,}"))}
final$pre.rating<-as.numeric(sapply(j,dig))
names(final)[1:3]<-c('name', 'state', 'total.points')
head(final)
```

```{r, eval=FALSE, echo=FALSE}
"'in the below, we try to arrive at selection for numbers until an alpha character is reached. Although our syntax works, we're still left with multiple elements if we include an '|' operator - necessary so that we capture those scores that do not include an alpha chharacter. We model the syntax after the one used on the initial step, where we force a stop at P (see above) '"

#([[:digit:]]){2,}P.+?(?=-)") - previous code from initial step
unlist(str_extract_all(j[41],"[[:digit:]]{2,}")) #works

dig<-function(x) {unlist(str_extract_all(x,"([[:digit:]]){2,}?(?=[A-Z])|[[:digit:]]{2,}"))}
dig(j[23])
dig(j[61])
w<-unlist(sapply(j,dig))
```

Calculate the average pre-chess rating of opponents. This requires a look-up of player number then a calculation.

Opponent ids are stored in columns [4:10] on the 'h.t' (raw) dataframe. I'll need to create a function that extracts the numeric portion from each of these fields which then looks up that player's pre-rating, and ultimately taking the mean from all of the opponents played. 

First do some data exploring to ensure that null values are not present.
I do this by creating a vector containing all of the opponents per player.
```{r}
#consider re-concatenating the columns for later application of regex
f<-str_c(h.t$V4,h.t$V5,h.t$V6,h.t$V7,h.t$V8,h.t$V9,h.t$V10,sep=" ")

#create function to extract digits
opp.f<-function(x){unlist(str_extract_all(x,"[[:digit:]]{1,}"))}
rt<-sapply(f,opp.f)#working with a list
tr<-lapply(rt,unlist)
v<-sapply(rt,length)
names(v)<-NULL
any(v!=6)
```

We see that indeed some players play more and less than others.
```{r}
table(v[v!=6])
#can index like this because of normalization
head(h.t[v!=6,4:10])
#ensure that no player has a null or 0 pre-rating
#anyNA(final[,4])
```

This data exploration informs us that not all players have played 6 rounds, and so the function we build must be able to handle this.
```{r}
#rt[1] #a list
#unlist rt, using lapply, do a lookup and calculate the mean
look<-function(x){ 
  k<-as.numeric(unlist(x))
  mean(final[k,4])
}
#names(d)<-NULL
final$pre.opp<-sapply(tr,look)
```


```{r, eval=FALSE, echo=FALSE}
#Name extraction function building:
xtract.nom<-function(x){
  nom<-unlist(str_extract_all(x,"[:alpha:]+ [:alpha:]+{2,}"))
  return(nom)
}
tre<-sapply(raw_split[4],xtract.nom)

unlist(str_extract_all(raw_split[4],"[[:alpha:]]+ [[:alpha:]]+"))
#pg 203
unlist(str_extract_all(raw_split[4],"( [[:alpha:]]+ ){2,}"))

test<-"| mama mia   | 45 | ma"
unlist(str_extract_all(test,"\\b| +[[:alpha:] ]{3,} +|\\b"))
unlist(str_extract_all(test,"\\|[ [:alpha:]]+ +\\|"))
unlist(str_extract_all(test,"\\| +[[:alpha:] ]{2,}"))
unlist(str_count(test, "\\| +\\w+ +\\|"))

unlist(str_extract_all(raw_split[3],"[[:alpha:]]{2,}"))


#"[[:alpha:]]
unlist(str_extract_all(j[3],"[[:alpha:]]{2,}"))
#assumes everyone only has two names
unlist(str_extract_all(j[3],"[[:alpha:]]+ [[:alpha:]]+"))
```

Before exporting to a csv, let's do a visualization of the data to look for outliers.

First, we'll plot a histogram of player pre-ratings... we'll then compare this to a histogram of average opponent pre-ratings.
```{r, warning=FALSE, message=FALSE}
#names(final)
ggplot(final, aes(final$pre.rating))+geom_histogram()+labs(title="Player's Pre-Rating Distribution")
```

```{r, warning=FALSE, message=FALSE}
ggplot(final, aes(final$pre.opp))+geom_histogram()+labs(title="Average Rating of Player Opponent")
```

Last thing... let's check that outlier player rating < 400...
```{r}
final[final$pre.rating==377,]
h.t[46,1:12]
```
I think we're good: the regex extracted ratings the same way the other players' ratings were. We confirmed using the sample on the project direction sheet.

Finally, write this out to a csv:
```{r, eval=FALSE}
write.csv(final,'your/directory/chess.csv')
```