---
title: "Assignment 2"
author: "Luis Calleja"
date: "September 17, 2016"
output: html_document
---

```{r}
library(stringr)
names <-"555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson,Homer5553642Dr. Julius Hibbert"
#Simpson,Homer5553642Dr. Julius Hibbert"
```

### Question #3a
Create a vector containing elements of names, then we rearrange the names in cases where last name appears first to arrive at a first name, last name orientation.
```{r}
#extract only names... the regex breaks/splits on non alpha or (.,,,_)
vec<-unlist(str_extract_all(names,"[[:alpha:]., ]{2,}"))

#in cases of commas, rearrange order of name
rearrange<- function(x) {
  if(!is.na(str_extract(x,","))){
    hol<-unlist(str_split(x,","))
    new<-str_c(hol[2],hol[1],sep=" ")
    return(new)
  }  else x
}

test<-sapply(vec,rearrange)

print(test)
```

### Question #3b
Construct a logical vector indicating whether "Rev." or "Dr." appears in the text.
I provide two ways to accomplish this.
```{r}
vec %in% vec[grep("(Rev.|Dr.)",vec)]
str_detect(vec,"(Rev.|Dr.)")
```

### Question #3c
Construct a vector indicating whether a character has a second name. I interpret this to be more than two names after removing for titles, eg. Rev., Dr.
```{r, eval=FALSE, echo=FALSE}
unlist(str_extract_all(test[6],"[^\bDr.\b]{2,}[[:alpha:]]{2,}"))

unlist(str_extract_all(test[3],"[^\b(Dr.|Rev.)\b]{2,}[[:alpha:]]{2,}"))
```

```{r}
mult.name<-function(x){
  g<-unlist(str_extract_all(x,"[^\b(Dr.|Rev.)\b][[:alpha:].]{2,}"))
  g<-str_trim(g,side="both")
  nom<-unlist(str_split(g, ' '))
  le<-length(nom)
  return(le)
}
#mult.name(test[2])
#unlist(str_extract_all(test[4],"[^\b(Dr.|Rev.)\b][[:alpha:]]{2,}"))
gh<-sapply(test,mult.name)
gh>2
```   

### Question #4a
[0-9]+\\$ describes a string that contains one ore more digits and ends with a dollar sign ("$"). We make use of the double forward slashes \ to force R to interpret the dollar sign to a literal. 
```{r}
unlist(str_extract_all("$99.948789$","[0-9]+\\$"))
unlist(str_extract_all("$99.99$","[0-9]+\\$"))
```
### Question #4b
the "\b" symbol analyzes the entire word; blank spaces are
treated as element breaks in this example.

The word must begin and end with a letter... notice how the first line does not return any matches. Forward slash b (\ b) forces whole word matching, we place another forward slash in front of it to serve as an escape character.
```{r}
unlist(str_extract_all("yoo7","\\b[a-z]{1,4}\\b"))
unlist(str_extract_all("yoot","\\b[a-z]{1,4}\\b"))

#unlist(str_extract_all("boo trip cam fgu7","\\b[a-z]{1,4}\\b"))
# compare without the requires that it end and start with a "b"
#unlist(str_extract_all("boob7oa","b[a-z]{1,4}b"))
```

### Question #4c
With ".*?\\.txt$", we are looking for a file that ends in a ".txt". the double forward slashes serve as an escape for the period, so that R interprets it as a literal. The "?" symbol allows for flexibility - removing the requirement of a character in front of the ".txt".
```{r}
unlist(str_extract_all("67call.txt",".*?\\.txt$"))
# asterisks = the preceding item will be 
#matched zero or more times
unlist(str_extract_all(".txt",".*?\\.txt$"))
```   

### Question #4d
"d" here is a symbol representing all digits [0-9]. Specifically, d{2} = any two digits - one following the other.
```{r}
# "\\d" = any digit
unlist(str_extract_all("99/45/19878","\\d{2}/\\d{2}/\\d{4}"))
#returns nothing, because the pattern is not sequential
unlist(str_extract_all("99jk/45/19878","\\d{2}/\\d{2}/\\d{4}"))
```

### Question #4e
Using syntax "<(.+?)>.+?</\\1>" we are looking for html elements. We ensure that each element is returned whole by invoking the callback feature - "\\1" ensuring that whatever text appeared in between the paranthesis at the beginning of the pattern, also appears at the end.
```{r, eval=FALSE, echo=FALSE}
unlist(str_extract_all("</","<(.+?)>.+?</\\1>"))
# "?" = at most once, but optional
unlist(str_extract_all("<(jj)>4","<(.+?)>.+?"))
#notice how this returns nothing
unlist(str_extract_all("<(jj)>4jog","<(.+?)> .+"))
#this will return the entire string: "<(jj)>4jog"
unlist(str_extract_all("<(jj)>4jog","<(.+?)>.*"))
#the paranthesis serve for framing a phrase we will 
#backreference later in the regex
unlist(str_extract_all("<jj>4jog","<(.+?)>.*"))
```

```{r}
unlist(str_extract_all("<head> Title Header </head>","<(.+?)>.+?</\\1>"))
```

```{r, echo=FALSE,eval=FALSE}
unlist(str_extract_all("<jj>4</j>","<(.+?)>.+?</\\1>"))
#appears as though the "?" symbol has a different impact if 
#placed at the end of the regex
unlist(str_extract_all("<head>Te</head>","<(.+?)>.+?"))
```
