---
title: "Assignment 1 - SQL and R"
author: "Luis Calleja"
date: "September 10, 2016"
output: html_document
---

In this exercise, we will create an R dataframe from a few tables of a Postgres database. The data describes viewer sentiment of movies they watch. 

Five movie goers were asked to provide a rating on six different movies - all of them from 2015. The movies were popular during the time in the box office - receiving recommendations from the New York Times or being in the top 50 of grossing movies for that year.

Several bits of information are recorded for both the critics and movies. Our final dataframe will contains all this information when we run our query.

### Postgres Retrieval

We are working with four tables:  
 - movie  
 + critic  
 * movie_genres  
 * movie_rating  
 
The dataset from the postgres db is retrieved in the form of a query (*"select"* statement).

First, we establish the connection to the database.
```{r, echo=FALSE}
pw='persyy'
```

```{r}
library(RPostgreSQL)
library(dplyr)
drv<- dbDriver("PostgreSQL")
con<-dbConnect(drv,dbname='ratings_movies', host='localhost',port=5432 , user='lechuza',password=pw)
```

With the connections made, let's extract all the data from all the tables and put it in a dataframe.
```{r}
w_df<-dbGetQuery(con,'select 
movie.title, critic.lastname, critic.firstname, movie_ratings.rating, movie_genres.genre 
from movie_ratings 
join critic on movie_ratings.critic_id_fkey=critic.critic_id
join movie on movie_ratings.movie_id_fkey=movie.movie_id
join movie_genres on movie_genres.movie_id_fkey = movie.movie_id 
order by movie.title, movie_genres.genre, lastname, firstname;')
```

Let's take a look at what we have.
```{r}
head(w_df,15)
```
Finito...