---
title: "Week 5 Assignment"
author: "Luis Calleja"
date: "September 30, 2016"
output: html_document
---
In this report we compare on-time performance between two carriers: AM West and Alaska Airlines. Our dataset is the ontime and delay frequency for multiple airports over the same sample period.
```{r, message=FALSE}
library(dplyr);library(tidyr);
library(ggplot2)
```

```{r}
raw<-read.csv('~/Documents/CUNY/data_class/week4-tidyData/flights.csv')
raw_tbl<-tbl_df(raw)
#raw_tbl
```

For efficient analysis, we will transform the data into a long format.
```{r}
#names(raw)
raw.c<-raw[complete.cases(raw),]
#raw_tbl[complete.cases(raw_tbl),]
```

We need to normalize the data and fill in missing values in the 'carrier' field. I'll show how to do this using base R and tidyr.

Base R:
```{r}
names(raw.c)[1:2]<-c('carrier','track')
#some prep work
g<-raw.c %>%
  gather(destination,freq,Los.Angeles:Seattle) 

g[g$carrier=='',1]<-rep(c('Alaska','AM WEST'),I(length(g$carrier)/4))

head(g)
```
Using tidyr functions:
```{r}
h<-raw.c %>%
  gather(destination,freq,Los.Angeles:Seattle) %>%
  mutate(carrier.c=as.character(carrier), carrier.1=ifelse(carrier.c=='',lag(carrier.c),carrier.c)) %>%
  select(carrier.1,track:freq)

head(h)
```

```{r, echo=FALSE,eval=FALSE}
length(g$carrier)/2
g[g$carrier=='',1]=rep(c('Alaska','AM WEST'),5)

h<-g %>%
  filter(carrier=='') %>%
  mutate(carrier2=rep(c('Alaska','AM WEST'),I(length(carrier)/2)))

h<-g %>%
  mutate(carrier2=ifelse(carrier!='',carrier,rep(c('Alaska','AM WEST'),I(length(carrier)/2))))

raw.c %>%
  gather(destination,freq,Los.Angeles:Seattle) %>%
  spread(carrier,c(destination,freq))

  mutate(carrier2=ifelse(carrier!='',carrier,rep(c('Alaska','AM WEST'),I(length(carrier)/2))))
```

### Final Step
Perform analysis to compare the arrival delays for the two airlines using dplyr and the grouping and summarise functions.

First statistic is the proportion of tardiness by carrier and city. Having this, we calculate average rank across all cities for the carriers (using the min_rank() function).

```{r, echo=FALSE,eval=FALSE}
tbl_df(g)
#mutate performs function on entire columns, unless we apply grouping
g %>%
  mutate(total=cumsum(freq))
vignette("window-functions")

#discover the minimum number of delays by carrier
g %>%
  mutate(min_rank=min_rank(freq))
```

```{r}
h.1<-h %>%
  group_by(carrier.1,destination) %>%
  mutate(total=sum(freq), prop=round(freq/total,2))

head(h.1)

h.1 %>%
  group_by(destination) %>%
  filter(track=='delayed') %>%
  mutate(rank=min_rank(prop)) %>%
  select(-c(freq:prop))
```
Alaska Airlines is the best performer (relative to its own activity) for each airport.

We prove this by looking at average rank across airport:
```{r}
h.1 %>%
  group_by(destination) %>%
  filter(track=='delayed') %>%
  mutate(rank=min_rank(prop)) %>%
  ungroup() %>%
  group_by(carrier.1) %>%
  summarise(avg.rank=mean(rank))
```
### Graphics
We run some charts to display volatility/distribution of delay proportions by carrier.
```{r}
#boxplot of proportional delays by destination
ggplot(h.1[h.1$track=='delayed',],aes(x=carrier.1,y=prop))+geom_boxplot()+ggtitle("Variance in delays by Carrier")+labs(x="Carrier",y="proportion of delays")
```   
The plots demonstrates how much more tardy AM West tends to be: 75% of AM Wests service has a higher proportion of delays than the typical delay at Alaska Airlines.

The below plot shows a destination-by-destination view of the proportion delayed. There is an overarching dominance by AM West.
```{r}
ggplot(h.1[h.1$track=='delayed',],aes(y=prop,x=factor(destination),fill=factor(carrier.1),color=factor(carrier.1)))+geom_bar(stat="identity",position="dodge")+theme(axis.text.x=element_text(angle=45, hjust=1))+ggtitle("Proportion of Delays by City")+labs(x="Destination",y="proportion of delays")
```

### Conclusion
Alaska performs better by way of proportion of flights delayed to on time arrival. This study readily lends itself to a Bayesian probability estimation model. With the model we would estimate the probability of being delayed given a destination airport and particular carrier.